---
title: "Personal Tracking Analysis - Insight Demo"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv")
head(data, 15)
str(data)
summary(data)

complete <- data[complete.cases(data), ]
head(complete, 15)
str(complete)
summary(complete)
```

## Average number of steps taken per day

```{r}
complete.sum <- tapply(complete$steps,complete$date,sum)
hist(complete.sum, breaks = 20, ylab = "Frequency", xlab = "Steps Per Day", main = "Histogram of Steps Per Day")
complete.mean <- mean(complete.sum[complete.cases(complete.sum)])  
```
The mean is: 
```{r, echo = FALSE}
complete.mean
```

```{r}
complete.median <- median(complete.sum[complete.cases(complete.sum)])
```
The median is: 
```{r, echo = FALSE}
complete.median
```

## Average Daily Activity Pattern

```{r}
time.series <- tapply(complete$steps,complete$interval,mean)
plot(y=time.series,x=row.names(time.series), type = 'l', xlab = "5-Minute Intervals", ylab ="Average Number of Steps Taken", main ="Time Series Plot of 5 Minute Intervals")

max <- which.max(time.series)
```
The 5 minute interval with the maximum average of steps on all days is:
```{r, echo = FALSE}
names(max)
```

## Imputing missing values

The total number of missing values is: 
```{r}
sum(is.na(data))
```

```{r}
newdata <- data
for (i in 1:nrow(newdata)) {
    if (is.na(newdata$steps[i])) {
        newdata$steps[i] <- time.series[which(newdata$interval[i] == row.names(time.series))]
    }
}
new.sum <- tapply(newdata$steps,newdata$date,sum)
hist(new.sum, breaks = 20, ylab = "Frequency", xlab = "Steps Per Day", main = "Histogram of Steps Per Day (with Imputed Values)")
new.mean <- mean(new.sum)
```

The mean of the imputed data is: 
```{r, echo=FALSE}
new.mean
```

```{r}
new.median <- median(new.sum)
```

The median of the imputed data is:
```{r, echo=FALSE}
new.median
```
The difference in the means are:
```{r, echo=FALSE}
abs(new.mean-complete.mean)
```
The difference in the medians are:
```{r, echo=FALSE}
abs(new.median-complete.median)
```

## Weekdays vs. Weekends

```{r}
newdata$weekdays <- ifelse(as.POSIXlt(newdata$date)$wday %in% c(0,6), 'weekend', 'weekday')
averagednewdata <- aggregate(steps ~ interval + weekdays, data=newdata, mean)
averagednewdataWeekend <- subset(averagednewdata, weekdays == "weekend")
averagednewdataWeekday <- subset(averagednewdata, weekdays == "weekday")
```

```{r, echo=FALSE, results='hide'}
#plot(y=averagednewdataWeekend$steps,x=averagednewdataWeekend$interval, type = 'l', xlab = "5-Minute Intervals", ylab = "Average Number of Steps Taken", main ="Time Series Plot of the Weekend", ylim=c(0,250))
summary(averagednewdataWeekend$steps)

#plot(y=averagednewdataWeekday$steps,x=averagednewdataWeekday$interval, type = 'l', xlab = "5-Minute Intervals", ylab = "Average Number of Steps Taken", main ="Time Series Plot of the Weekdays", ylim=c(0,250))
summary(averagednewdataWeekday$steps)
```

```{r}
linewidth <- 2

plot(c(0,2500),c(0,250),type= 'n', xlab= "5-Minute Intervals", ylab = "Average Number of Steps Taken")
lines(x = averagednewdataWeekend$interval, y = averagednewdataWeekend$steps,col="red",lwd=linewidth) 
lines(x = averagednewdataWeekday$interval, y = averagednewdataWeekday$steps,col="blue",lwd=linewidth) 
legend(1900, 250, c("Weekday", "Weekend"), lty=c(1,1), lwd=c(linewidth,linewidth), col=c("blue","red")) 
summary(averagednewdataWeekend$steps)
summary(averagednewdataWeekday$steps)
```
